---
title: "Analyse MALDI-MSI ovary dataset"
output:
  html_document:
  keep_md: true
smart: false
toc: true
toc_float: true
theme: united
params:
  yaml_file: yaml_file
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***
  
# Setup
## Libraries
```{r}
#load R packages
library(Seurat)
library(dplyr)
library(ggplot2)
library(yaml)
```

## Parameters from YAML
```{r}
args <- yaml.load_file(params$yaml_file)
```

# Cluster pixels and visualize on UMAP
## Cell clustering
```{r}
#load MALDI-MSI data exported from SCILS
pixels <- read.csv(args$input_files$pixels)
masses <- read.csv(args$input_files$masses)
metadata <- read.csv(args$input_files$metadata)

colnames(pixels) <- c('pixelID', masses$x)
rownames(pixels) <- paste0("pixel_", pixels[,1])
pixels <- pixels[,-1]
```

## Explore Raw data
```{r}
dim(pixels)
pixels[1:5,1:5]
pixels[1:5,438:443]
```
## Use m/z > 400
```{r}
n_pcs <- args$analysis_params$n_pcs
res <- args$analysis_params$resolution

msidata.data <- pixels[,as.numeric(colnames(pixels))>400]

# msidata.data <- read.csv(file='1.csv',row.names = 1, header = TRUE,sep=",")
#transform of data format
msidata.data <- msidata.data*100 # Why x100?
msidata.data <- round(msidata.data,digits=0)
msidata.data <- as.data.frame(t(msidata.data)) # Transpose, so pixels become columns
#create Seurate object
msidata <- CreateSeuratObject(counts = msidata.data, project = "IRI")
# data normalization
msidata <- SCTransform(msidata, verbose = FALSE)
# PCA analysis
msidata <- RunPCA(msidata, assay = "SCT", verbose = FALSE)
VizDimLoadings(msidata, dims = 1:2, reduction = "pca")
DimPlot(msidata, reduction = "pca")
DimHeatmap(msidata, dims = 1:15, cells = 500, balanced = TRUE)
ElbowPlot(msidata)

# UMAP analysis
msidata <- FindNeighbors(msidata, dims = 1:n_pcs)
msidata <- FindClusters(msidata, resolution = res)

# Notice the parameters diverting from default. Is this needed in our case?
msidata <- RunUMAP(object = msidata,
                   dims = 1:n_pcs)# ,
                   # n.neighbors = 15L,
                   # min.dist = 0.05,
                   # check_duplicates = FALSE)
# saveRDS(object = msidata,
#         file = "msidata.rds")
DimPlot(object = msidata, reduction = 'umap', label = TRUE, group.by="orig.ident",pt.size=1,label.size=5)
DimPlot(object = msidata, reduction = 'umap', label = TRUE, pt.size=1,label.size=5)
```

## Add cluster id from Seurat object to the metadata
```{r}
seurat_clusters <- msidata@meta.data %>% select(seurat_clusters)
seurat_clusters$pixelID <- rownames(seurat_clusters)
# seurat_clusters$seurat_clusters <- paste0('cl_', seurat_clusters$seurat_clusters)
metadata$pixelID <- paste0('pixel_', metadata$pixelID)
metadata <- inner_join(metadata, seurat_clusters, by="pixelID")
# saveRDS(object = metadata, file = 'metadata.rds')
```

## Plot with ggplot2 and geom_raster
```{r}
p <- ggplot(metadata, aes(x, y)) +
  geom_raster(aes(fill=seurat_clusters)) +
  labs(title = "Seurat clusters in space")
print(p)
# ggplotly(p)
```

## Select features
* Features with highest residual variance. These are the most variable features
* For more details, refer to: https://htmlpreview.github.io/?https://github.com/satijalab/sctransform/blob/supp_html/supplement/variance_stabilizing_transformation.html
```{r}
sct_feature_attributes <- msidata@assays$SCT@SCTModel.list$model1@feature.attributes
sct_feature_attributes %>% arrange(desc(residual_variance)) %>% head(n=20)
top <- sct_feature_attributes %>% arrange(desc(residual_variance)) %>% head(n=20) %>% rownames()
```

## Plot top most variable metabolites
```{r}
# Plot as pct of top value for each metabolite (column)
pixels_top_int <- pixels[ ,top]
column_max <- pixels_top_int %>% summarise_if(is.numeric, max)
pixels_top_int_pct <- as.data.frame(round(mapply('/', pixels_top_int, column_max)*100))
pixels_top_int_pct$pixelID <-  row.names(pixels_top_int)
intensities_pct <- merge(metadata,
                         pixels_top_int_pct,
                         by = "pixelID")
intensities_pct <- intensities_pct %>% select(-c(pixelID, dataset, normalization, seurat_clusters))
for (feature in top) {
  print(feature)
  p <- ggplot(intensities_pct, aes(x, y)) +
    geom_raster(aes(fill=get(feature))) +
    scale_fill_gradient(low = "black", high = "green") +
    labs(title = paste0("Intensities for m/z of: ", feature))
  print(p)
}
```
## Plot metabolites with m/z > 400
```{r}
top <- colnames(pixels)[as.numeric(colnames(pixels))>400]

# Plot as pct of top value for each metabolite (column)
pixels_top_int <- pixels[ ,top]
column_max <- pixels_top_int %>% summarise_if(is.numeric, max)
pixels_top_int_pct <- as.data.frame(round(mapply('/', pixels_top_int, column_max)*100))
pixels_top_int_pct$pixelID <-  row.names(pixels_top_int)
intensities_pct <- merge(metadata,
                         pixels_top_int_pct,
                         by = "pixelID")
intensities_pct <- intensities_pct %>% select(-c(pixelID, dataset, normalization, seurat_clusters))
for (feature in top) {
  print(feature)
  p <- ggplot(intensities_pct, aes(x, y)) +
    geom_raster(aes(fill=get(feature))) +
    scale_fill_gradient(low = "black", high = "green") +
    labs(title = paste0("Intensities for m/z of: ", feature))
  print(p)
}
```


